---
title: "CEM1002"
author: "Neil Montgomery"
date: "2015-09-14"
output: beamer_presentation
---
```{r echo=FALSE, include=FALSE, message=FALSE}
## Import any libraries and set global options here. 
library(ggplot2)
library(dplyr)
```


## Course Basics

- Neil Montgomery
- neilm@mie.utoronto.ca
- BA8132 office hours...any time

- Official description is misleading (from 2013)
- Course will be primarily "statistics" with cities issues as applications
- I'm working on field trips and/or guest lectures
- Evaluation:
    * 4 assignments each worth 25%
    * An assignment may have multiple parts with different due dates (esp. the first one)
- No required texts. For the basics, any recent "Stats 101" book is fine.
    
## Software

Statistics requires software. In this course we'll use R. 

Pros:

- High quality free software environment
- Help desk: Google
- Extensions via rich package eco-system
- Not point-and-click (reproducible workflow)

Cons:

- Not point-and-click (learning curve)

To install:

- Optional but recommended: Install \LaTeX\ (to produce PDF reports) `latex-project.org`
- Install `R` itself: `r-project.org`
- Install Rstudio: `rstudio.com`

## Assignment 1 (full details to be on course website)

- Find a dataset related to one or more cities and perform a detailed exploratory data analysis (EDA) on it.
- Hint: Google "<CITY NAME> open data"
- Three deadlines:
    * September ~~21~~ **22**: Install `R` environment, find (interim) dataset(s), and produce a skeleton report with a numerical summary and a graphical summary.
    * September ~~28~~ **29**: Using final dataset(s) provide interim report.
    * October ~~5~~ **9**: Final report.
    * Dates may be adjusted to avoid distaster.

## R miscellanea

- Notes and R code used in each class will be available.
- A few packages will get used quite a bit and you'll need to install them (once) and import them (every time). Start with:
    * `dplyr`
    * `ggplot2`
- Rstudio makes good workflow easy with R markdown. Examples:
    * Two plus three equals `r 2+3`.
    * The `diamonds` example dataset that comes with `ggplot2` has `r nrow(diamonds)` records in it.
    
## R miscellanea (plot example)

- The plotting package `ggplot2` will seem strange at first but is the gold standard.

```{r echo=FALSE, fig.height=4, fig.width=6}
# qplot(carat, price, data = diamonds)
qplot(log(carat), log(price), color = color, data = diamonds)
```

## Exploratory Data Analysis

- What is a dataset?
- Classification of variable types (not to be taken too seriously)
- Numerical summaries
- Graphical summaries 


